Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:
```{r loading packages}
library(dplyr)
library(ggplot2)
```

```{r Third Qualitative Variable}
# pf <- read.delim('C:\\Users\\Ziyad\\Dropbox\\Nano Degree\\4\\4eda-course-materials\\lesson3\\pseudo_facebook.tsv')
pf.fc_by_age_gender <- pf %>% 
  filter (!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count), 
            median_friend_count = mean(friend_count), 
            n= n()) %>%
  ungroup() %>%
  ungroup(age)
tail(pf.fc_by_age_gender)  
library(ggplot2)

```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot (data= pf.fc_by_age_gender,  aes(x= age,y= median_friend_count)) +
  geom_line(aes(col= gender))

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
install.packages('tidyr')
library(tidyr)
pf.fc_by_age_gender.wide <-     subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],                   !is.na(gender)) %>% 
    spread(gender, median_friend_count) %>% 
    mutate(ratio = male / female)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot (data= pf.fc_by_age_gender.wide,  aes(x= age,y= female / male)) +
  geom_line()+
  geom_hline(yintercept = 1, lty= 2, alpha = 0.3)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004,2009, 2011, 2012,2014))
table(year_joined.bucket)


```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot (subset(pf, !is.na(year_joined.bucket)), aes(x= age, y= friend_count))+
  geom_line(aes(col= year_joined.bucket), stat = 'summary', fun.args= median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot (subset(pf, !is.na(year_joined.bucket)), aes(x= age, y= friend_count))+
  geom_line(aes(col= year_joined.bucket), stat = 'summary', fun.args= mean) +
  geom_line(stat = 'summary', fun.args = mean, lty= 2)

```

***

### Friending Rate
Notes:

```{r Friending Rate}
with (subset(pf, tenure >0 ), summary(friend_count/ tenure))

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot (data= subset(pf, tenure >0 ), aes(x=tenure ,y= friendships_initiated / tenure)) +
  geom_line(aes(col= year_joined.bucket))
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.args = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.args = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.args = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.args = mean)

library(gridExtra)
grid.arrange()
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

yo$id <- factor(yo$id)
str(yo)

ggplot (yo, aes(x=price))+
  geom_histogram()
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yo$all.purchases <- yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry
#the below code line is equla to the above code line
# yo <- transform(yo, all.purchases =  strawberry + blueberry + pina.colada + plain + mixed.berry)

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.id <- sample (levels(yo$id), 16)

ggplot (data= subset(yo, id%in% sample.id), aes(x= time,y= price))+
  facet_wrap(~id)+
  geom_line()+
  geom_point(aes(size= all.purchases), pch= 1)
  
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r Scatterplot Matrix}
#the below code will take a long ime around 1 hour to run 
'''
install.packages("GGally")
library(GGally)
theme_set(theme_minimal(20))


set.seed(1836)
pf_subset <- pf[,c(2:15)]
names(pf_subset)
ggpairs (pf_[sample.int(pf_subset), 1000],)
'''
```


***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

