Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
pf <- read.delim('pseudo_facebook.tsv') 

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
install.packages('ggthemes', dependencies = TRUE) 
library(ggthemes)
library(ggplot2)

names(pf)
qplot (x = dob_day ,data = pf ) +scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
ggplot (pf, aes(x= dob_day)) +
  geom_histogram(binwidth = 1)+
  scale_x_continuous(breaks =1:31)+
  facet_wrap(~dob_month)

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
ggplot (pf, aes(x = friend_count)) +
  geom_histogram()

```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
ggplot (pf, aes(x= friend_count)) +
  geom_histogram() +
  scale_x_continuous(limits = c(1,1000))

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram( binwidth = 10)+
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram( binwidth = 10)+
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_grid(gender~.)

```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
ggplot (data = subset(pf, !is.na(gender)), aes(x= friend_count)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = seq(1,1000,50), limits = c(1,1000)) +
  facet_wrap(~gender)
  

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table (pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
ggplot(pf, aes(x = tenure)) +
  geom_histogram(binwidth = 30, fill = '#099DD9', col='black')



```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
ggplot(aes(x = tenure/365), data = pf) + 
   geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
ggplot(aes(x = tenure/365), data = pf) + 
   geom_histogram(binwidth = .25, color = 'black', fill = '#F79420') +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')

```

***

### User Ages
Notes:

```{r User Ages}
ggplot (pf, aes(x= age)) +
  geom_histogram(col='black', fill= 'red',   limits= c(1:120)) +
  scale_x_continuous(breaks =seq(0,120,10))

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
install.packages('gridExtra') 
library(gridExtra)

p1 <- ggplot (pf, aes(x= friend_count)) +
  geom_histogram()
  
p2 <- p1 + scale_x_log10()

p3 <- p1 + scale_x_sqrt()

grid.arrange(p1,p2,p3,ncol=1)
```

***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
by (pf$www_likes, pf$gender, sum)
ggplot (subset(pf, !is.na(gender)), aes(x= www_likes)) +
  geom_histogram(aes(fill = gender)) +
  scale_x_continuous() + 
  scale_x_log10()

```


***

### Box Plots
Notes:

```{r Box Plots}
ggplot (data = subset(pf,!is.na(gender)), aes(x=gender , y= friend_count)) +
  geom_boxplot()

```


#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
ggplot (data = subset(pf,!is.na(gender)), aes(x=gender , y= friend_count)) +
  geom_boxplot() +
  scale_y_continuous(limits= c(1,1000))
by(pf$friend_count, pf$gender, summary)

# equals to the following 

ggplot (data = subset(pf,!is.na(gender)), aes(x=gender , y= friend_count)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(1,1000))
by(pf$friend_count, pf$gender, summary)
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
by (pf$friendships_initiated, pf$gender, summary)


```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
ggplot(subset(pf, !is.na(gender)), aes(x=gender ,y=friendships_initiated) ) +
  geom_boxplot() +
  coord_cartesian(ylim = c(1,125))

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary (pf$mobile_likes)
summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes >0 ,1 ,0 )
pf$mobile_check_in <- factor(pf$mobile_check_in)

summary(pf$mobile_check_in)
sum(pf$mobile_check_in)/ length(as.numeric(pf$mobile_check_in))
```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!