Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
install.packages('ggplot2')
library('ggplot2')
pf <- read.delim('C:\\Users\\Analysis-PC04\\Dropbox\\Nano Degree\\4\\4eda-course-materials\\lesson3\\pseudo_facebook.tsv')

ggplot(pf, aes(x= age, y= friend_count)) +
  geom_point()
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(pf, aes(x= age, y= friend_count)) +
  geom_point()
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(pf, aes(x= age, y= friend_count)) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(pf, aes(x= age, y= friend_count)) +
  geom_point(alpha = 1/20, position = position_jitter(height = 0)) +
  xlim(13, 90) +
  coord_trans(y= 'sqrt')

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(pf, aes(x= age, y= friend_count)) +
  geom_point(alpha = 1/20, position = position_jitter(height = 0)) +
  xlim(13, 90) +
  coord_trans(y= 'sqrt')

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot (pf, aes(x= age, y= friendships_initiated)) +
  geom_point()

ggplot (pf, aes(x= age, y= friendships_initiated)) +
  geom_point()

ggplot (pf, aes(x= age, y= friendships_initiated)) +
  geom_point(alpha= 1/10, position = position_jitter(height = 0))

ggplot (pf, aes(x= age, y= friendships_initiated)) +
  geom_point(alpha= 1/10, position = position_jitter(height = 0))+
  coord_trans(y='sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
#install.packages('dplyr')
#library('dplyr')
#library('ggplot2')
pf.fc_by_age <- group_by (pf, age) %>%
  summarise(friend_count_mean = mean(friend_count),friend_count_median= median(friend_count),n = n()) %>%
  arrange(age)
head(pf.fc_by_age)
 
ggplot(pf.fc_by_age, aes(x=age ,y=friend_count_mean)) +
  geom_line()

```

Create your plot!

```{r Conditional Means Plot}
ggplot(pf, aes(age, friend_count)) +
  xlim(13, 90) +
  geom_point(position = position_jitter(height = 0),
              col = 'orange',
              alpha = 0.05) +
  coord_trans(y='sqrt')+  
  geom_line(stat = 'summary',  fun.y = mean)+
  geom_line(stat = 'summary', fun.y = "quantile", fun.args = list(probs =0.1 ), col='blue', lty=2)+
  geom_line(stat = 'summary', fun.y = "quantile", fun.args = list(probs =0.5 ), col='blue')+
  geom_line(stat = 'summary', fun.y = "quantile", fun.args = list(probs =0.9 ), col='blue', lty=2 )

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(pf, aes(age, friend_count)) +
  geom_point(position = position_jitter(height = 0),
              col = 'orange',
              alpha = 0.05) +
  coord_cartesian(xlim = c(13,70), ylim = c(9,1000))+
  geom_line(stat = 'summary',  fun.y = mean)+
  geom_line(stat = 'summary', fun.y = "quantile", fun.args = list(probs =0.1 ), col='blue', lty=2)+
  geom_line(stat = 'summary', fun.y = "quantile", fun.args = list(probs =0.5 ), col='blue')+
  geom_line(stat = 'summary', fun.y = "quantile", fun.args = list(probs =0.9 ), col='blue', lty=2 )
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
age_friend <- pf %>% select(age, friend_count)
with(pf , cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age >= 70) , cor.test(age, friend_count, method = 'pearson'))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot (pf, aes(x=www_likes_received, y= likes_received)) +

summary(pf$www_likes_received)
summary(pf$likes_received)

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot (pf, aes(x=www_likes_received, y= likes_received)) +
  geom_point() +
  #this line belo DOES NOT WORK
  #coord_cartesian(xlim = c(0,quantile(pf$www_likes_received, 0.95)), ylim = quantile(pf$likes_received, 0.95))
  xlim(0,quantile(pf$www_likes_received, 0.95))+
  ylim(0,quantile  (pf$likes_received    , 0.95))+
  geom_smooth(method = 'lm', col= 'red')
summary(pf$www_likes_received)
summary(pf$likes_received)

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with (pf , cor.test(www_likes_received, likes_received, method = 'pearson'))

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
#library(alr3)
data("Mitchell")
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot (Mitchell, aes(x= Month, y=Temp )) +
  geom_point()

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot (Mitchell, aes(x= Month, y=Temp )) +
  geom_point()+
  scale_x_continuous(breaks = seq(0, 203,12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
# pf$age_with_months <- 
pf$age_with_months <- pf$age + (1 - pf$dob_month / 12) 

```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(
  friend_count_mean= mean(friend_count),
  friend_count_median= median(friend_count), 
  n= n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}


ggplot (data= subset(pf.fc_by_age_months, age_with_months < 70), aes (x= age_with_months, y= friend_count_mean)) +
  geom_line()

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
p1 <- ggplot (data= subset(pf.fc_by_age, age <= 70), aes (x= age, y= friend_count_mean)) +
  geom_line()+
  geom_smooth()

p2 <- ggplot (data= subset(pf.fc_by_age_months, age_with_months <= 70), aes (x= age_with_months, y= friend_count_mean)) +
  geom_line() +
  geom_smooth()

p3 <- ggplot (data= subset(pf, age <= 70), aes (x= round(age / 5)* 5 , y= friend_count)) +
  geom_line(stat = 'summary', fun.args= mean)

# install.packages('gridExtra')
# library('gridExtra')
grid.arrange(p2, p1, p3, ncol=1)
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

